Self-tunable approximated explicit MPC: Heat exchanger implementation and analysys

### 2023-02-02 paper/elsarticle/main.tex
results:energy consumption results

	The control performance was also investigated quantitatively. Table~\ref{tab:control_performance} summarizes the evaluated control performance criteria computed for the two boundary controllers and for the self-tuned controller. The control performance is evaluated for each reference step change separately. The considered quality criteria are: sum-of-squared based criterion -- integral squared error $ISE$, maximal overshoot/undershoot $\sigma_{\mathrm{max}}$, settling time $t_{\epsilon}$ for 5\,\%-neighbourhood of the reference temperature $T_{\mathrm{ref}}$, and the volume of the hot medium $V$ consumed for the corresponding control. To provide better readability of the computed results in Table~\ref{tab:control_performance}, the best values, i.e., the minimum values, are emphasized using a bold font style.

	\begin{table}[h!]
		\begin{center}
			\caption{Control performance criteria.}
			\label{tab:control_performance}
			\begin{tabular}{c|c|c|c|c|c} 
				Reference step change & $Q_\mathrm{y}$ & $ISE$ [$^{\circ}\mathrm{C}^2$\,s] & $\sigma_{\mathrm{max}}$\,[\%] & $t_{\epsilon}$\,[s] & $V$\,[l] \\
				\hline
				\multirow{2}{*}{ 35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 714 & 33.50 & 16.5 & \textbf{2.12} \\
				& 100 & 867 & 16.65 & 12.5 & 2.36 \\ 
				& self-tuned & \textbf{678} & \textbf{15.19} & \textbf{9.5} & 2.38 \\ 
				\hline
				\multirow{2}{*}{ 45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C } & 1000 & 365 & 47.20 & \textbf{5} & \textbf{2.49} \\
				& 100 & 606 & 23.25 & 26.5 & 3.19 \\ 
				& self-tuned & \textbf{248} & \textbf{19.13} & 9.5 & 3.35 \\ 
				\hline
				\multirow{2}{*}{ 50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 245 & \textbf{18.92} & \textbf{6.5} & \textbf{2.00} \\
				& 100 & 398 & 79.64 & 31 & \textbf{2.00} \\ 
				& self-tuned & \textbf{186} & 24.59 & \textbf{6.5} & 2.10 \\ 
				\hline
				\multirow{2}{*}{ 45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C } & 1000 & 1024 & 18.43 & 22.5 & 0.94 \\
				& 100 & 1402 & 41.87 & 90 & \textbf{0.93} \\ 
				& self-tuned & \textbf{967} & \textbf{16.49} & \textbf{18.5} & 1.10  
			\end{tabular}
		\end{center}
	\end{table}

	As can be seen in Table~\ref{tab:control_performance}, that real-time self-tuning of the explicit MPC controller helped to improve two to three criteria when tracking each reference value. The cost for this is the energy associated with control. Although the integral square error, maximal overshoot/undershoot and settling time decreased, the volume of consumed heating medium did not. Nevertheless, the average deterioration in the terms of consumed heating medium is approximately by~17\,\%.

@@

	\begin{table}[h!]
		\begin{center}
			\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
			\label{tab:improvement}
			\begin{tabular}{c|c|c|c|c} 
				Reference step change & $\delta ISE$\,[\%] & $\delta \sigma_{\mathrm{max}}$\,[\%] & $\delta t_{\epsilon}$\,[\%] & $\delta V$\,[\%] \\
				\hline
				35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C &  5 &  9 & 24 &$-12$ \\ 
				% \hline
				45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 32 & 18 &$-90$ & $-35$ \\ 
				% \hline
				50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C & 24 &$-30$& 0 &$-4$\\ 
				% \hline
				45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C &  6 & 11 & 18 &$-18$  
			\end{tabular}
		\end{center}
	\end{table}
	%\begin{table}[h!]
		%\begin{center}
			%\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
			%\label{tab:improvement}
			%\begin{tabular}{c|c|c|c|c} 
				%Reference step change & $\Delta ISE$\,[\%] & $\Delta \sigma_{\mathrm{max}}$\,[\%] & $\Delta t_{\epsilon}$\,[\%] & $\Delta V$\,[\%] \\
				%\hline
				%35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C & 5.04 & 8.77 & 24.00 & -12.26 \\ 
				%\hline
				%45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 32.05 & 17.72 & -90.00 & -34.54 \\ 
				%\hline
				%45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 24.08 & -29.97 & 0 & -4.48 \\ 
				%\hline
				%45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C & 5.57 & 10.53 & 17.78 & -18.28  
				%\end{tabular}
			%\end{center}
		%\end{table}

	Implementing a self-tunable explicit MPC controller leads to improved control performance in the most analyzed quality criteria, see Table~\ref{tab:control_performance}. In general, utilizing a proposed controller with a scalable aggressiveness according to the operating conditions leads to higher accuracy (lower $ISE$), lower value of the overshoots (lower $\sigma_{\mathrm{max}}$), and faster achieving the reference value (lower $t_{\epsilon}$). %It should be also noted that the deterioration of every criterium is calculated subject to the best out of \textit{two} other controllers. It follows that the evaluation is strict as in practice only one controller would be available.
	Although the volume of the hot medium $V$ is not reduced in the considered control scenario, the improved control performance may lead to other benefits.

	\textcolor{red}{[ TODO: How about to simply omit the evaluation of the volume $V$ and write something around these lines: ]}

@@

%	\begin{center}
		\begin{tabular}{ l l }
			$u_{\mathrm{L}}$ & control inputs associated with the lower boundary controller\\
			$u_{\mathrm{U}}$ & control inputs associated with the upper boundary controller\\
			$U$ & voltage, \% \\
			$U^{\mathrm{s}}$ & steady state of voltage, \% \\
			$\mathcal{U}$ & set of control inputs \\
			$V$ & volume of heating medium, l \\
			$x$ & system states \\
			$\widetilde{x}$ & augmented system states \\
			$x_{\mathrm{I}}$ & system states corresponding to the built-in integrator \\
			$y$ & system outputs \\
			$y_\mathrm{\max}$ & maximal value of system outputs \\
			$y_\mathrm{\min}$ & minimal value of system outputs \\
			$y_\mathrm{ref}$ & reference value of system outputs \\
			$\mathcal{Y}$ & set of system outputs \\
			\textit{0} & zero matrix
		\end{tabular}
%	\end{center}

>>>>

	The control performance was also investigated quantitatively. Table~\ref{tab:control_performance} summarizes the evaluated control performance criteria computed for the two boundary controllers and for the self-tuned controller. The control performance is evaluated for each reference step change separately. The considered quality criteria are: sum-of-squared based criterion -- integral squared error $ISE$, maximal overshoot/undershoot $\sigma_{\mathrm{max}}$ and the settling time $t_{\epsilon}$ for 5\,\%-neighbourhood of the reference temperature $T_{\mathrm{ref}}$. To provide better readability of the computed results in Table~\ref{tab:control_performance}, the best values, i.e., the minimum values, are emphasized using a bold font style.

%	\begin{table}[h!]
%		\begin{center}
%			\caption{Control performance criteria.}
%			\label{tab:control_performance}
%			\begin{tabular}{c|c|c|c|c|c} 
%				Reference step change & $Q_\mathrm{y}$ & $ISE$ [$^{\circ}\mathrm{C}^2$\,s] & $\sigma_{\mathrm{max}}$\,[\%] & $t_{\epsilon}$\,[s] & $V$\,[l] \\
%				\hline
%				\multirow{2}{*}{ 35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 714 & 33.50 & 16.5 & \textbf{2.12} \\
%				& 100 & 867 & 16.65 & 12.5 & 2.36 \\ 
%				& self-tuned & \textbf{678} & \textbf{15.19} & \textbf{9.5} & 2.38 \\ 
%				\hline
%				\multirow{2}{*}{ 45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C } & 1000 & 365 & 47.20 & \textbf{5} & \textbf{2.49} \\
%				& 100 & 606 & 23.25 & 26.5 & 3.19 \\ 
%				& self-tuned & \textbf{248} & \textbf{19.13} & 9.5 & 3.35 \\ 
%				\hline
%				\multirow{2}{*}{ 50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 245 & \textbf{18.92} & \textbf{6.5} & \textbf{2.00} \\
%				& 100 & 398 & 79.64 & 31 & \textbf{2.00} \\ 
%				& self-tuned & \textbf{186} & 24.59 & \textbf{6.5} & 2.10 \\ 
%				\hline
%				\multirow{2}{*}{ 45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C } & 1000 & 1024 & 18.43 & 22.5 & 0.94 \\
%				& 100 & 1402 & 41.87 & 90 & \textbf{0.93} \\ 
%				& self-tuned & \textbf{967} & \textbf{16.49} & \textbf{18.5} & 1.10  
%			\end{tabular}
%		\end{center}
%	\end{table}

\begin{table}[h!]
	\begin{center}
		\caption{Control performance criteria.}
		\label{tab:control_performance}
		\begin{tabular}{c|c|c|c|c} 
			Reference step change & $Q_\mathrm{y}$ & $ISE$ [$^{\circ}\mathrm{C}^2$\,s] & $\sigma_{\mathrm{max}}$\,[\%] & $t_{\epsilon}$\,[s]  \\
			\hline
			\multirow{2}{*}{ 35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 714 & 33.50 & 16.5 \\
			& 100 & 867 & 16.65 & 12.5 \\ 
			& self-tuned & \textbf{678} & \textbf{15.19} & \textbf{9.5}  \\ 
			\hline
			\multirow{2}{*}{ 45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C } & 1000 & 365 & 47.20 & \textbf{5} \\
			& 100 & 606 & 23.25 & 26.5  \\ 
			& self-tuned & \textbf{248} & \textbf{19.13} & 9.5  \\ 
			\hline
			\multirow{2}{*}{ 50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 245 & \textbf{18.92} & \textbf{6.5}  \\
			& 100 & 398 & 79.64 & 31  \\ 
			& self-tuned & \textbf{186} & 24.59 & \textbf{6.5}  \\ 
			\hline
			\multirow{2}{*}{ 45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C } & 1000 & 1024 & 18.43 & 22.5  \\
			& 100 & 1402 & 41.87 & 90  \\ 
			& self-tuned & \textbf{967} & \textbf{16.49} & \textbf{18.5}   
		\end{tabular}
	\end{center}
\end{table}

	As can be seen in Table~\ref{tab:control_performance}, that real-time self-tuning of the explicit MPC controller helped to improve two to three criteria when tracking each reference value. The cost for this is the energy associated with control.

@@

	\begin{table}[h!]
		\begin{center}
			\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
			\label{tab:improvement}
			\begin{tabular}{c|c|c|c} 
				Reference step change & $\delta ISE$\,[\%] & $\delta \sigma_{\mathrm{max}}$\,[\%] & $\delta t_{\epsilon}$\,[\%]  \\
				\hline
				35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C &  5 &  9 & 24 \\ 
				% \hline
				45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 32 & 18 &$-90$  \\ 
				% \hline
				50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C & 24 &$-30$& 0 \\ 
				% \hline
				45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C &  6 & 11 & 18   
			\end{tabular}
		\end{center}
	\end{table}
	%\begin{table}[h!]
		%\begin{center}
			%\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
			%\label{tab:improvement}
			%\begin{tabular}{c|c|c|c|c} 
				%Reference step change & $\Delta ISE$\,[\%] & $\Delta \sigma_{\mathrm{max}}$\,[\%] & $\Delta t_{\epsilon}$\,[\%] & $\Delta V$\,[\%] \\
				%\hline
				%35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C & 5.04 & 8.77 & 24.00 & -12.26 \\ 
				%\hline
				%45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 32.05 & 17.72 & -90.00 & -34.54 \\ 
				%\hline
				%45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 24.08 & -29.97 & 0 & -4.48 \\ 
				%\hline
				%45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C & 5.57 & 10.53 & 17.78 & -18.28  
				%\end{tabular}
			%\end{center}
		%\end{table}

		\begin{table}[h!]
			\begin{center}
				\caption{Energy consumption evaluation.}
				\label{tab:energy}
				\begin{tabular}{c|c} 
					$Q_\mathrm{y}$ & E\,[kJ]   \\
					\hline
					1000 &  883  \\ 
					% \hline
					100 & 938   \\ 
					% \hline
					self-tuned & 921   
				\end{tabular}
			\end{center}
		\end{table}

	Implementing a self-tunable explicit MPC controller leads to improved control performance in the most analyzed quality criteria, see Table~\ref{tab:control_performance} and Table~\ref{tab:improvement}. In general, utilizing a proposed controller with a scalable aggressiveness according to the operating conditions leads to higher accuracy (lower $ISE$), lower value of the overshoots (lower $\sigma_{\mathrm{max}}$), and faster achieving the reference value (lower $t_{\epsilon}$). %It should be also noted that the deterioration of every criterium is calculated subject to the best out of \textit{two} other controllers. It follows that the evaluation is strict as in practice only one controller would be available.

@@

%	\begin{center}
		\begin{tabular}{ l l }
			$u_{\mathrm{L}}$ & control inputs associated with the lower boundary controller\\
			$u_{\mathrm{U}}$ & control inputs associated with the upper boundary controller\\
			$U$ & voltage, \% \\
			$U^{\mathrm{s}}$ & steady state of voltage, \% \\
			$\mathcal{U}$ & set of control inputs \\			
			$x$ & system states \\
			$\widetilde{x}$ & augmented system states \\
			$x_{\mathrm{I}}$ & system states corresponding to the built-in integrator \\
			$y$ & system outputs \\
			$y_\mathrm{\max}$ & maximal value of system outputs \\
			$y_\mathrm{\min}$ & minimal value of system outputs \\
			$y_\mathrm{ref}$ & reference value of system outputs \\
			$\mathcal{Y}$ & set of system outputs \\
			\textit{0} & zero matrix
		\end{tabular}
%	\end{center}
